The low dimensional world of echolocating bats

### 2020-03-25 Paper/main.tex
after grammarly

We have suggested before that limitations of the bat's sonar system result in bats having access to a sparse, low dimensional description of the outside 3D world only \citep{Vanderelst2015a,Vanderelst2016}. Some recent evidence seems to confirm this. First, \citet{Geberl2019} confirmed that bio-sonar has a low spatial resolution, both for distance and angular dimensions. Also, neurophysiological data suggest that closely spaced objects are not perceived seperatly by echolocating bats. In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018,Knowles2015}.

However, it should be noted that such low dimensional descriptions have been shown to potentially contain a lot of useful information \citep{Kuc1997b,Kuc1997}. We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}. Under this hypothesis, bats compare the acoustic signature of the environment to a set of previously stored signatures. We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.

@@

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of uncorrelated features, irrespective of whether we use artificial or real echo sequences. Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, could also support flight control.

@@

We created artificial echo sequences by randomly spacing ideal point reflectors. The distances of the point reflectors were selected using a two-step process. First we selected $n_{s}$ seed points $s$, uniformly distributed over a range from \MinDistance\ to \MaxDistance\ meters. Next, for every seed point $s$, we selected $n_{c}$ cloud points normally distributed around $s$, with standard deviation 0.25 m. We added a single reflector at a distance between \MinDistanceFirst\ and \MaxDistanceFirst\ to each collection of points. This two step process to modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points. We generated \GenerateRepeats\ sets of points for each combination of $n_s \in \ns$ and $n_c \in \nc$. Synthetic echoes consisting of few reflectors are likely to consist of only weak echoes.

Each set of points was converted into an environmental impulse response with a sampling frequency $f_s$ of 219 kHz. The environmental impulse response length was set to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In representing each point reflector by a Dirac impulse, we modeled attenuation of the echoes by introducing an amplitude calculated based on spherical spreading (both ways). The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB. This final operation represents a very simple model of varying reflector strength due e.g., to shape and orientation of the reflecting feature.

@@

\begin{figure*}[tb]
	\centering
	\includegraphics[width=1\linewidth]{../results/processing}
	\caption{Example of synthetic echo sequence and its processing using the \citet{Wiegrebe2008} model. (a) Example of a single synthetic echo sequence (red). The dechirped version is plotted in black. (b-i) Response of the \nfilters\ frequency channels in the \citet{Wiegrebe2008} model. From these plots, it can be seen that channels correlate highly. This is confirmed in panel j, depicting the average correlation across channels for all synthetic echo sequences analyzed. Note that the x-scale does not start at zero. (k) The output of the \citet{Wiegrebe2008} model summed across the frequency channels, is referred to as a template in this paper. Panel (l) shows the cumulative proportion of explained variance as a function of number of \pc\ across all templates for all synthetic echo sequences.}
	\label{fig:processing}
\end{figure*}

@@

In this section, we  used a generative model for naturalistic echoes. We assumed that echoes sequences received by bats can be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted into low dimensional descriptions. The original echo sequences, generated by \MinDiracs\ to \MaxDiracs\ point reflectors effectively map to a \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

We reused the large library of echoes gathered in real bat habitats by \citet{Vanderelst2016}. Below, we briefly discuss the construction of templates based on these echoes, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}. The emitter was a Sensecomp 7000 Series ultrasonic broadband speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation. The device and pan-tilt system was mounted on a tripod.

@@

As we did for the synthetic echo sequences, the echo trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated. Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile, or template. As the beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012} templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using the emissions themselves for analysis. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

@@

The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes, described above. We assessed how well the templates collected in real habitats can be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space. Next, we calculated the correlation (and explained variance) between the original templates and the reconstructed templates.

@@

We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates can be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table. The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation requires a comprehensive search of memory that scales with the number of templates memorized. Hence, in our previous work, we established the theoretical possibility of classifying templates based on real echoes from environments bats were known to frequent, irrespective of the significant amount of memory and computational resources required to do so.  Here, we use a more biologically plausible memory system: an artificial neural network, i.e. a neurally inspired way of distributed information storage \citep{Mcleod1998}. While we acknowledge that current neural network training methods most likely have no biological counterpart, we propose that this approach provides us with a more biologically plausible appraisal of the memory and computational effort required to implement a similar scheme in a bat brain.

Simple artificial feedforward neural networks were implemented using Tensorflow 2 for Python. The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$). The following layers were fully connected and used the rectified linear activation function. The output layer uses a softmax activation function, i.e. the normalized exponential function. The loss was calculated as the cross-entropy between the target output and the actual output of the neural network. Weights were updated after each epoch using the Adam algorithm \citep{Kingma2014} .

@@

We should point out that we did not take any steps to avoid so-called overfitting by the neural networks. Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract complicated relationships between inputs and outputs that do not hold beyond the training examples. Overfitting typically reduces the ability of the model to generalize its predictive power to new data. However, in this case, the model is not required to generalize to other environments. Instead, we are only interested in whether the neural networks can learn and classify the set of training examples. Nevertheless, it would be beneficial if the neural networks would be capable of classifying templates corresponding to directions that are slightly different than those included in the data sets. Indeed, this is a feature of the Euclidean distance measure used in implementing the perfect memory. In that approach, novel templates will be classified to positions with similar templates. We verified whether this properties was also found in the neural networks the case by generating a set of templates by interpolating in azimuth and elevation between the directions ensonified in the data set. In our data sets, we ensonified 31 different azimuth directions and 7 elevation directions. The interpolated data consisted of 30 azimuths and 7 elevations. We used the neural networks trained on the original data to classify these interpolated data.

@@

Interestingly, for azimuth and elevation, classification errors were not uniform across directions. This can be seen by the classification entropy plotted in figure \ref{fig:israelperformance}. In particular, these curves show  peaks around zero azimuth and medium elevations. The geometry of the environment can explain these peaks. The corridor (fig. \ref{fig:datacollection}) does not return a lot of echoes from azimuths around zero. Indeed, this direction is mostly open space. The same is true for elevations above 0.

@@

Figure \ref{fig:royalperformance} shows the results for the Royal Fort Park data. These results mirror those in figure \ref{fig:israelperformance}. Again the neural network performed at least as good as the perfect memory approach. For azimuth, the perfect memory approach slightly outperform the neural network (notice that y-scale in the top panels of this figure differ from figure \ref{fig:israelperformance}). Overall, the performance for this data set was better than for the Israel data. This is probably due to the fact that echoes were collected closer to reflectors (vegetation) than in the Israel data set, which leads to strong echoes.

@@

We assumed echoes received by bats can be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}). This suggests that the filtering of the auditory periphery provides the bat with a significantly compressed acoustic description of its environment. To identify the locus of the compression, we performed a number of additional analyses. In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters and/or the low-pass filter in the \citet{Wiegrebe2008} model.

We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model; with and without the 1 Khz low-pass filter. Per default, the model includes this filter to model the response characteristics of the inner hair cells \citep{Meddis2006,Reijniers2010a}. By removing this lowpass filter, we can assess to what extent it is responsible for the observed compressibility of the templates. We also generated templates using echo sequences of which the samples were randomly shuffled. This allows inferring whether the low-dimensionality of the templates is a property of the echo signals (before being filtered by the cochlear model).

Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance. From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from \pca\ to \FigVIIInumberIII\. Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to \FigVIIInumberII. Including the 1 kHz filter, the average spectrum of the templates (fig. \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz. Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kSamples/sec. This corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate. This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig. \ref{fig:mechanism}). The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca. Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction. The structure of the echo signals also decreases the dimensionality of the templates. The sonar system of bats, through a number of filtering stages, compresses the 3D complexity of the world onto a lower dimensional representation.

@@

While recognizing location in the environment is a critical component of navigation, bats use their sonar not only for navigating. They also rely on sonar for flight control. This includes, avoiding obstacles and/or following guiding structures. The environmental description the bat has access to for flight control starts from the same cochlear output  used for place recognition. Therefore, our current work suggests that the same low-dimensional description should be sufficient for flight control as well. Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. In that work, we converted both the echoes at the left and at the right ear into a template-like representation used as input to the motion controller. Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One could presume a neural network that converts the \pca\ features to flight control parameters.

@@

Despite their small size, the networks in this paper were able to encode (memorize) the templates at least as well as the perfect memory approach. The neural networks presented here are tiny compared to the density of cortical matter. However, the networks are also small compared to the data they encode. Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates. The number of floating point numbers making up the templates is larger than the number of weights by a factor of \Compression. Therefore, the networks have a smaller memory footprint than the data they encode. It should be noted that we have not made an attempt to try and to optimize (minimize) the size of the networks. From these numbers, we conclude that the low-dimensional signals provided by the bat's sonar system result in a low memory footprint and that they would use little neural resources to be stored and processed.

\begin{table}
	\centering
\begin{tabular}{lrrrrl}
	\toprule
	Site &       T &     n &      W &  W (nb) &      R \\
	\midrule
	Israel &  183430 &  5395 &  52875 &   51975 &  28.8\% \\
	Royal &  264146 &  7769 &  51345 &   50475 &  19.4\% \\
	\bottomrule
\end{tabular}
	\label{tab:sizes}
	\caption{Table indicating the size of both template data sets and number of weights in the networks. The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set. The number of weights $W$ includes all connections between neurons as well as the bias weights for each of the three networks trained. $W (nb)$ omits the bias weights. The ratio $R$ gives the ratio $W/T$.}
\end{table}

@@

It has been suggested before that bats' sonar systems provides detailed information about the world. A number of authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}. However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar. In particular, we argued before that the neurophysiology of the  bats auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat. Indeed, \citet{Reijniers2010a} already noticed that significant compression occurs during cochlear transduction. Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of peripheral auditory processing.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{../results/interpolation}
	\caption{Neural network categorization results for the templates obtained by interpolating between the 7 elevation and 31 azimuth directions ensonified for the Israel and Royal Fort site. (a,b) results for the azimuth and elevation for the Israel data. (c,d) Same for the Royal Fort data. Because the templates classified in these results were interpolated, the network can be considered to have correctly classified the template if it is assigned to direction $n$ or direction $n+1$. For example, a template obtained by interpolating between azimuth directions 140 and 130 degrees could be assigned to either these directions. The number in each panel indicates the percentage of data point on the diagonal (direction $n$) or the diagonal + 1 ($n+1$).}
	\label{fig:interpolation}
\end{figure}

>>>>

We have suggested before that limitations of the bat's sonar system result in bats having access to a sparse, low dimensional description of the outside 3D world only \citep{Vanderelst2015a,Vanderelst2016}. Some recent evidence seems to confirm this. First, \citet{Geberl2019} confirmed that bio-sonar has a low spatial resolution, both for distance and angular dimensions. Also, neurophysiological data suggest that closely spaced objects are not perceived separately by echolocating bats. In contrast, neural responses to echo cascades from densely spaced objects seem to represent a single extended stimulus \citep{Warnecke2018,Knowles2015}.

However, it should be noted that such low dimensional descriptions have been shown to contain a lot of useful information \citep{Kuc1997b,Kuc1997} potentially. We previously demonstrated that bats might use sonar template-based scene recognition \citep{Vanderelst2016,Vanderelst2017}. Under this hypothesis, bats compare the acoustic signature of the environment to a set of previously stored signatures. We operationalized these acoustic signatures as time-intensity profiles, which could be derived directly from the cochlear output.

@@

Together these results suggest that echolocating bats live in a low dimensional perceptual world. The physics of the sonar system and the auditory periphery map the complexity of the 3D environment onto a small number of uncorrelated features, irrespective of whether we use artificial or real echo sequences. Nevertheless, this limited set of features is sufficiently rich for bats to recognize locations, and as we discuss, they could also support flight control.

@@

We created artificial echo sequences by randomly spacing ideal point reflectors. The distances of the point reflectors were selected using a two-step process. First we selected $n_{s}$ seed points $s$, uniformly distributed over a range from \MinDistance\ to \MaxDistance\ meters. Next, for every seed point $s$, we selected $n_{c}$ cloud points normally distributed around $s$, with standard deviation 0.25 m. We added a single reflector at a distance between \MinDistanceFirst\ and \MaxDistanceFirst\ to each collection of points. This two-step process of modeling natural reflectors was inspired by the results of \citet{Yovel2009}, who found that echoes returning from vegetation could be better modeled using a similar two-step process than simply using uniformly spaced points. We generated \GenerateRepeats\ sets of points for each combination of $n_s \in \ns$ and $n_c \in \nc$. Synthetic echoes consisting of few reflectors are likely to consist of only weak echoes.

Each set of points was converted into an environmental impulse response with a sampling frequency $f_s$ of 219 kHz. The environmental impulse response length was set to 7499 samples. These parameters correspond to the recording length and the sampling frequency used in the acoustic data collection (see below). In representing each point reflector by a Dirac impulse, we modeled attenuation of the echoes by introducing an amplitude calculated based on spherical spreading (both ways). The final amplitude of each Dirac impulse was determined by sampling from a normal distribution centered on the calculated amplitude and with a standard deviation of 10 dB. This last operation represents a straightforward model of varying reflector strength due, e.g., to shape and orientation of the reflecting feature.

@@

\begin{figure*}[tb]
	\centering
	\includegraphics[width=1\linewidth]{../results/processing}
	\caption{Example of synthetic echo sequence and its processing using the \citet{Wiegrebe2008} model. (a) Example of a single synthetic echo sequence (red). The dechirped version is plotted in black. (b-i) Response of the \nfilters\ frequency channels in the \citet{Wiegrebe2008} model. From these plots, it can be seen that channels correlate highly. This is confirmed in panel j, depicting the average correlation across channels for all synthetic echo sequences analyzed. Note that the x-scale does not start at zero. (k) The output of the \citet{Wiegrebe2008} model summed across the frequency channels. This is referred to as a template in this paper. Panel (l) shows the cumulative proportion of explained variance as a function of the number of \pc\ across all templates for all synthetic echo sequences.}
	\label{fig:processing}
\end{figure*}

@@

In this section, we used a generative model for naturalistic echoes. We assumed that echoes sequences received by bats could be considered as generated by anisotropically distributed point reflectors. Using this model, we have derived a low dimensional space ($D = \pca$), which can be used to describe the templates resulting from the artificial echo sequences. This implies that, after filtering through (a model of) the bat's auditory system, these artificial echoes can be converted into low dimensional descriptions. The original echo sequences, generated by \MinDiracs\ to \MaxDiracs\ point reflectors effectively map to a \pca\ dimensional vector. In the remainder of the paper, we will assess whether this low dimensional space allows describing the echoes from real bat habitats as collected using an ensonification device.

@@

We reused the extensive library of echoes gathered in real bat habitats by \citet{Vanderelst2016}. Below, we briefly discuss the construction of templates based on these echoes, and we refer to our previous work for more details. \citet{Vanderelst2016} constructed an ensonification device featuring 31 Knowles FG series microphones connected to a custom-built sonar data acquisition board \citep{Steckel2013a}. The emitter was a Sensecomp 7000 Series ultrasonic broadband speaker. The ensonification device was mounted on a pan-tilt system allowing to rotate it in azimuth and elevation. The device and pan-tilt system was mounted on a tripod.

@@

As we did for the synthetic echo sequences, the echo trains collected in bat habitats were dechirped and filtered using a model of the bat's auditory periphery as proposed by \citet{Wiegrebe2008}. We concluded from analyzing the synthetic echoes that the response in different frequency channels is highly correlated. Therefore, the output of the cochlear model was averaged across frequency channels to obtain a time-intensity profile or template. As the beam of the emitter used was more narrow than the typically combined hearing and emission directionality of bats \citep{Vanderelst2010a,Reijniers2010,Jakobsen2012} templates were averaged across three neighboring directions. The first 6 ms of data were removed to avoid using the emissions themselves for analysis. The processing steps outlined above resulted in 3 $\times$ 217 templates, for each location at each site, corresponding to 217 directions and three repeats.

@@

The templates from both sites were mapped into the \pca-dimensional space derived from the synthetic echoes, described above. We assessed how well templates collected in real habitats could be represented by the \pca\ components by mapping the templates into this space and transforming the results back to the original space. Next, we calculated the correlation (and explained variance) between the original templates and the reconstructed templates.

@@

We assessed whether the templates projected into the \pca-dimensional space still contained sufficient information to be recognized. Previously, we assessed whether templates could be used to recognize locations using, what \citet{Baddeley2012} called, a perfect memory approach. Templates were stored in a lookup table. The Euclidean distance between two echo templates was the metric used to assess how well bats should be able to distinguish between templates. This table lookup and distance calculation require a comprehensive search of memory that scales with the number of templates memorized. Hence, in our previous work, we established the theoretical possibility of classifying templates based on real echoes from environments bats were known to frequent, irrespective of the significant amount of memory and computational resources required to do so.  Here, we use a more biologically plausible memory system: an artificial neural network, i.e., a neurally inspired way of distributed information storage \citep{Mcleod1998}. While we acknowledge that current neural network training methods most likely have no biological counterpart, we propose that this approach provides us with a more biologically plausible appraisal of the memory and computational effort required to implement a similar scheme in a bat brain.

Simple artificial feedforward neural networks were implemented using Tensorflow 2 for Python. The first layer of the neural network added Gaussian noise to each input sample ($\sigma(\stochasticnoise) = 0.01$). The following layers were fully connected and used the rectified linear activation function. The output layer uses a softmax activation function, i.e., the normalized exponential function. The loss was calculated as the cross-entropy between the target output and the actual output of the neural network. Weights were updated after each epoch using the Adam algorithm \citep{Kingma2014} .

@@

We should point out that we did not take any steps to avoid so-called overfitting by the neural networks. Overfitting occurs when a neural network \citep{Ghotra2017}, or other mathematical models \citep{Hawkins2004}, extract complicated relationships between inputs and outputs that do not hold beyond the training examples. Overfitting typically reduces the ability of the model to generalize its predictive power to new data. However, in this case, the model is not required to generalize to other environments. Instead, we are only interested in whether the neural networks can learn and classify the set of training examples. Nevertheless, it would be beneficial if the neural networks would be capable of classifying templates corresponding to directions that are slightly different from those included in the data sets. Indeed, this is a feature of the Euclidean distance measure used in implementing the perfect memory. In that approach, novel templates will be classified as positions associated with similar templates. We verified whether this property was also found in the neural networks the case by generating a set of templates by interpolating in azimuth and elevation between the directions ensonified in the data set. In our data sets, we ensonified 31 different azimuth directions and seven elevation directions. The interpolated data consisted of 30 azimuths and six elevations. We used the neural networks trained on the original data to classify these interpolated data.

@@

Interestingly, for azimuth and elevation, classification errors were not uniform across directions. This can be seen by the classification entropy plotted in figure \ref{fig:israelperformance}. In particular, these curves show peaks around zero azimuth and medium elevations. The geometry of the environment can explain these peaks. The corridor (fig. \ref{fig:datacollection}) does not return many echoes from azimuths around zero. Indeed, this direction is mostly open space. The same is true for elevations above 0.

@@

Figure \ref{fig:royalperformance} shows the results for the Royal Fort Park data. These results mirror those in figure \ref{fig:israelperformance}. Again the neural network performed at least as good as the perfect memory approach. For azimuth, the perfect memory approach outperforms the neural network (notice that the y-scale in the top panels of this figure differ from figure \ref{fig:israelperformance}). Overall, the performance for this data set was better than for the Israel data. Probably, this is because echoes were collected closer to reflectors (vegetation) than in the Israel data set, which leads to strong echoes.

@@

We assumed echoes received by bats could be modeled as originating from a large collection of point reflectors. Passing these synthetic echoes through a model of the auditory periphery, we obtained a time-intensity profile (template) for each echo sequence. These templates could be well described by a low dimensional space of \pca\ principal components (figure \ref{fig:synthetictemplates}). We found that the same \pca D space could be used to describe templates originating from echoes collected in real bat habitats (figure \ref{fig:empiricaltemplates}). This suggests that the filtering of the auditory periphery provides the bat with a significantly compressed acoustic description of its environment. To identify the locus of the compression, we performed additional analyses. In particular, we evaluated whether the low-dimensional description is a property of the echo sequences themselves, or whether this property results from the finite bandwidths of the gammatone filters or the low-pass filter in the \citet{Wiegrebe2008} model.

We generated templates from the synthetic echo sequences using two versions of the \citet{Wiegrebe2008} model, with and without the 1 kHz low-pass filter. Per default, the model includes this filter to model the response characteristics of the inner hair cells \citep{Meddis2006,Reijniers2010a}. By removing this low-pass filter, we can assess to what extent it is responsible for the observed compressibility of the templates. We also generated templates using echo sequences of which the samples were randomly shuffled. This allows testing whether the templates' low-dimensionality is a property of the echo signals (before being filtered by the cochlear model).

Figure \ref{fig:mechanism} shows the number of \pcs\ needed to capture 99\% of the templates variance. From this plot, it can be seen that the number of \pcs\ increases when shuffling the echo sequences from \pca\ to \FigVIIInumberIII\. Omitting the 1 kHz filter from the  \citet{Wiegrebe2008} model increases the number of \pcs\ to \FigVIIInumberII. Including the 1 kHz filter, the average spectrum of the templates (fig. \ref{fig:mechanism}, inset) shows that most energy in the templates is contained in frequencies below 1 kHz. Therefore, if we take 1 kHz as the highest relevant frequency, it ought to be possible to completely represent templates by sampling them at 2 kSamples/sec. This corresponds to about 70 samples per template, as templates consist of 7499 samples at 219 kHz sample rate. This corresponds roughly to the number of \pcs\ required to represent templates derived from echo sequences that were shuffled  (fig. \ref{fig:mechanism}). The number of \pcs\ required to represent unshuffled templates is about half of this, i.c., \pca. Therefore, we conclude that the 1 kHz filter is the main, but not exclusive, locus of the dimensionality reduction. The structure of the echo signals also decreases the dimensionality of the templates. The sonar system of bats, through several filtering stages, compresses the 3D complexity of the world onto a lower-dimensional representation.

@@

While recognizing location in the environment is a critical component of navigation, bats use their sonar not only for navigating. They also rely on sonar for flight control. This includes avoiding obstacles and following guiding structures. The environmental description the bat has access to for flight control starts from the same cochlear output used for place recognition. Therefore, our current work suggests that the same low-dimensional description should be sufficient for flight control as well. Indeed, we have previously demonstrated that a low-dimensional description is sufficient for obstacle avoidance and corridor following \citep{Vanderelst2015a,Mansour2019}. In that work, we converted the echoes at the left and right ear into a template-like representation. We used this representation as input to the robot controller. Subsequent work will have to demonstrate whether the parameters of the \pca D space presented here can be used directly for controlling flight. In other words, we conjecture it should be possible to extract \pca\ features from an echo signal and convert this vector into a motor command. One could presume a neural network that converts the \pca\ features to flight control parameters.

@@

Despite their small size, the neural networks were able to encode (memorize) the templates at least as well as the perfect memory approach. The neural networks presented here are tiny compared to the density of cortical matter. However, the networks are also small compared to the data they encode. Table \ref{tab:sizes} lists the number of weights for the artificial neural networks and compares this number with the size of the \pca-dimensional templates. The number of floating-point numbers making up the templates is larger than the number of weights by a factor of \Compression. Therefore, the networks have a smaller memory footprint than the data they encode. It should be noted that we have not attempted to try and to optimize (minimize) the size of the networks. From these numbers, we conclude that the low-dimensional signals provided by the bat's sonar system result in a low memory footprint and that they would use limited neural resources to be stored and processed.

\begin{table}
	\centering
	\begin{tabular}{lrrrrl}
		\toprule
		Site &       T &     n &      W &  W (nb) &      R \\
		\midrule
		Israel &  183430 &  5395 &  52875 &   51975 &  28.8\% \\
		Royal &  264146 &  7769 &  51345 &   50475 &  19.4\% \\
		\bottomrule
	\end{tabular}
	\label{tab:sizes}
	\caption{Table indicating the size of both template data sets and the number of weights in the networks. The template size $T$ is obtained as $\pca \times n$, with $n$ the number of templates in the set. The number of weights $W$ includes all connections between neurons and the bias weights. $W (nb)$ omits the bias weights. The ratio $R$ gives the ratio $W/T$.}
\end{table}

@@

It has been suggested that bats' sonar systems provide detailed information about the world. Several authors have referred to sonar as providing the bat with an acoustic image \citep{Lee2017,Barchi2013,Moss2001,Schnitzler2003,Simmons2012,Ulanovsky2008,Clare2015,Surlykke2016,Geipel2013}. As mentioned above, this is at odds with some recent behavioral and neurophysiological findings in bats faced with complex echo sequences \citep{Knowles2015,Geberl2013,Warnecke2018}. However, it is also at odds with the various filtering stages involved in the emission and the reception of bio-sonar. In particular, we argued before that the neurophysiology of the bats' auditory system represents a bottleneck in the amount of raw echo signal information accessible to the bat. Indeed, \citet{Reijniers2010a} already noticed that significant compression occurs during cochlear transduction. Here, we followed up on that work to quantify the dimensional reduction of the input signals as a result of peripheral auditory processing.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=1\linewidth]{../results/interpolation}
	\caption{Neural network categorization results for the templates obtained by interpolating between the seven elevation and 31 azimuth directions ensonified for the Israel and Royal Fort site. (a,b) results for the azimuth and elevation for the Israel data. (c,d) Same for the Royal Fort data. Because the templates classified in these results were interpolated, the network can be considered to have correctly classified the template if it is assigned to direction $n$ or direction $n+1$. For example, a template obtained by interpolating between azimuth directions 140 and 130 degrees could be assigned to either these directions. The number in each panel indicates the percentage of data point on the diagonal (direction $n$) or the diagonal + 1 ($n+1$).}
	\label{fig:interpolation}
\end{figure}
